Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution

### 2017-04-02 paper.tex
intro revise

%
%===============================================================================
\section{Introduction}
%===============================================================================
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-RTL and post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (IP), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers/firmware) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or firmware).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing.

In the post-RTL phase, when the RTL design is available, hardware/software
co-verification becomes more complex.  Unlike the TLM, which only captures
limited design functionality, the RTL implements cycle-accurate behavior
of the final HW, and addresses concerns related to power, area and
timing.  Verification is mandatory to assert that the driver or firmware (FW) 
is executed correctly together with the RTL.  However, the verification of
FW given in C/C++ together with the RTL given in a hardware
description language (HDL) such as Verilog/VHDL is a challenging task due to
the timing and semantic mismatch between the two models -- for example,
FW running on a processor could be much faster than the HW
model it interacts with.  The effectiveness of testing is
thus severely limited in post-RTL co-verification.

Common practice in industry for system-level co-verification is to 
either use emulators/accelerators or Instruction Set Simulators (ISS). 
However, no rigorous formal verification effort is
performed at the post-RTL phase to ensure the validity of the
FW-HW interactions~\cite{vlsid13}.  With this aim in mind, the
primary contribution of this paper is to develop an automated and scalable
hardware/software co-verification methodology that detects FW-HW bugs in the 
presence of an RTL implementation.

A firmware only exercise some fragment of hardware state-space in a typical 
hardware/software co-design~\cite{dac17,taa,ibm-other}.  However, there is no 
automatic hardware/software co-verification tool that can exploit this interaction 
behavior.  For building effective and scalable formal co-verification tool, it is 
therefore imperative to automatically mine only relevant scenario/transaction pairs 
that occurs in practical co-designs.  In this paper, we attempt to bridge this gap by 
presenting a technique that automatically mine relevant HW/SW interaction behavior 
using symbolic execution.  A firmware services the requests from
the application software layer or its interacting hardware~\footnote{Note that whenever 
we refer to ``hardware'' from this point onwards, we mean an RTL implementation 
and not a TLM.}  components. 
Similarly, the hardware services the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in \emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the UART in \emph{transmitter} mode which restricts the communication between the FW and HW to set of services that only transfers data from the FW to the serial output port of the transmitter HW while completely bypassing the receiver HW.    A {\em Scenario/Transaction}
pair is a set of valid interactions between $\mathcal{HT}$ and
$\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable and effective co-verification.

A conventional, monolithic approach, e.g., an
approach based on BMC~\cite{CKY03}, cannot exploit this infeasibility.  The
key idea in this paper is to use path-based symbolic execution to
automatically infer such Scenario/Transaction pairs and check their validity
using backend solvers.

\Omit{
A system is comprised of a set of concurrent FW and HW transactions which
interleaves asynchronously. It is well recognized that concurrency is at
the heart of co-verification. The number of interleavings between the FW
and HW threads can be large, which can become a bottleneck in the
verification process. }

The concurrency between the HW and SW components is a key challenge in
co-verification.  In this paper, we exploit specific interaction patterns
between the FW and HW to reduce the number of interleavings.  We observe
that this interaction pattern follows a producer-consumer relationship --
the master module (producer) solely updates the shared registers which are
consumed by the slave (consumer).  This key observation allows us to
sequentialize a concurrent FW-HW model following specific interaction
patterns.

Generating a unified FW-HW model for co-verification is a well-known 
technique. Notably, Kurshan et al.~modeled HW and 
SW using finite state machines~\cite{fmsd02}, Monniaux in~\cite{emsoft07} modeled
HW and SW as C programs which are formally pushdown systems (PDS), 
Li et al.~in~\cite{fase10} used Buchi Automata to abstractly
model a hardware and PDS to abstractly model a software to generate a unified SW-HW model, 
called Buchi Pushdown System ({\em BPDS}). In this paper, we propose an 
orthogonal approach for co-verification.  
The HW RTL model in Verilog is automatically synthesized to a 
word-level cycle-accurate representation, called, 
{\em software netlist} $(\mathcal{SN})$, in ANSI-C~\cite{mtk2016}. 
The FW model is then instrumented with the call to the 
top level procedure of $\mathcal{SN}$ following specific FW-HW interaction 
patterns. Note that the top level procedure in $\mathcal{SN}$
is the same as the top level module in HW, since the 
synthesis step from RTL to $\mathcal{SN}$ preserves the 
RTL module hierarchy. This results in a unified single-threaded software model, 
that contains both FW and $\mathcal{SN}$ 
(synthesized hardware RTL). We call this 
unified FW-HW model, a {\em program netlist} (${\mathcal{PN}}$). 
This representation enables us to apply
broad range of verification techniques which have 
been less commonly explored in formal hardware/software 
co-verification based on netlists. In particular, we 
apply path-based symbolic execution and bounded model checking (BMC). 
In this paper, we present a path-based symbolic execution tool, 
\verifox, to uncover FW-HW design bugs in ${\mathcal{PN}}$. 
We compare our tool with the state-of-the-art BMC based 
co-verification tool, \hwcbmcv. Our experimental results show 
that path-based symbolic execution is order of magnitude 
faster than the BMC based approaches.

@@

\end{enumerate}

\section{Scenarios and Transaction}
%\section{Preliminaries}
In this section, we start by presenting a running example of FW, followed
by some definitions and notations.\\    
%===============================================================================
\subsection{Running Example}
{\bf Running example: }
%===============================================================================
Figure~\ref{firmware} gives a fragment of a FW model for a Universal 
Asynchronous Transmitter/ Receiver (UART) IP.  The
$wb$ class of functions communicate with the {\em wishbone} bus interface to
set/reset the UART input ports and read/write data through the bus
interface.  The FW implements Linux style inb() and outb() functions which
invokes the $wb$ functions to communicate with the UART.  Due to space
limitation, we could not accommodate all $wb$ functions.  The FW begins by
resetting the UART IP which is followed by a $wb\_idle()$ function.  The 
FW then configures the UART in {\em loopback} mode using $outb$ function calls. 
Inside the loop, the firmware checks for a {\em transmitter} or {\em
receiver} interrupt by reading the Interrupt Status register $(UART\_IS)$. 
Depending on the type of interrupt, the firmware either receives a data packet 
or sends a data packet over the serial output. We verify that the transmitted 
data is the same as the received data in the {\em loopback} mode using an 
assertion in the firmware model. Note that $set\_inputs()$ and
$next\_timeframe()$ functions, used in $wb$ functions, are 
primitives in \textsc{HW-CBMC} that direct the tool to set the 
inputs to the HW signals and advance the HW clock, respectively.

\Omit{
The $set\_inputs()$ and the $next\_timeframe()$ functions are primitives
in \textsc{HW-CBMC} that direct the tool to set the inputs to the HW
signals and advance the HW clock, respectively.  Recall that $\mathcal{SN}$
is the synthesized word-level representation of HW RTL. In the unified
FW-HW model, $\mathcal{PN}$, these primitives are implemented by invoking
the top level module of $\mathcal{SN}$ which is the top module of the UART
RTL IP core.
}

\Omit{ 
In Figure~\ref{firmware}, the firmware executes scenario, $\mathcal{S}$, 
to configure the UART IP in {\em loopback} mode. A property $\psi$ can be as follows: \\
{\em If the content of memory mapped registers (mode configuration and interrupt 
enable) is, $UART\_MC = 0x13$ and $UART\_IE = 0x03$,  then the serial output of 
the transmitter, $UART\_TX$, should not be in high-impedance $(Z)$ state and 
the transmitter buffer must not be EMPTY.}  \\
Note that the antecedent of the property (configuring the UART in 
loopback with interrupt enabled) defines a scenario and the 
consequent defines the condition on the HW transaction. 
In this example, $\mathcal{T}$ \sjcmt{Where did
$\mathcal{T}$ come from?} is a finite-duration service 
through the UART state machine, $(UART\_TX and UART\_RX)$, that 
is spread over multiple clock cycles. 
}

\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l}
\hline
Wishbone Interface & Main Module \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
typedef unsigned char u8;
unsigned char inb 
 (unsigned long port) {
  return wb_read(port);
}
void outb (u8 value, 
 unsigned long port) {
  wb_write(port, value);
}
void wb_reset(void) {
 rst_i = 1;
 set_inputs();
 next_timeframe();
 rst_i = 0;
 stb_i = 0; 
 cyc_i = 0;
}
void wb_idle() {
 set_inputs();
 next_timeframe();
}
void wb_write(_u32 addr, 
     _u8 b) {
 adr_i = addr;
 dat_i = b;
 we_i = 1;
 cyc_i = 1;
 stb_i = 1;
 set_inputs();
 next_timeframe();
 we_i = 0;
 cyc_i = 0;
 stb_i = 0;
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
int main() {
wb_reset();
wb_idle();
// Configure the uart
outb (0x13, UART_MC);  
outb (0x80, UART_CM3);
outb (0x00, UART_CM2);
outb (0x00, UART_CM1);
outb (0x00, UART_CR);  
outb (0x03, UART_IE); 
char tx_b[] = "Hello world";
_u8 status = 0;
char rx_b[100];
int i=0,c=0,d=0;
// data transfer in loopback
for (i=0; i<1990; i++){
if (irq_o){
 status=inb(UART_IS)&0x0c;
 if(istatus==0x0c){
 //it was a tx_empty interrupt
 outb(*(tx_b+c),UART_TR); c++;
 }
 else{ //status==0x04
 //it was an rx_data interrupt
  rx_b[d] = inb(UART_TR); d++;
 }
}else {
 // no interrupt. 
 wb_idle();
 wb_idle();
 }
}
// property
for(i=0; i<=10; i++)
 assert(rx_b[i] == tx_b[i]);
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Firmware model of UART IP}
\label{firmware} 
\end{figure}
%
\textbf{Scenario ($\mathcal{S}$):}
Let $B$ denote a set of all possible behaviors of a system. A scenario is a set of 
constraints $\mathcal{S}$ that restrict the system behaviors
to $B' \subseteq B$. For example, the left hand side of Figure~\ref{fig:st}. 
gives an example scenario for the UART FW shown in Figure~\ref{firmware}. 
Here, $\mathcal{S}$ configure the UART IP in {\em loopback} mode with 
interrupt enabled.

@@

Each execution of $\mathcal{HT}$ is a non-empty 
sequence of transaction steps where each transaction step 
happens in a single clock. The first step is called {\em initiation event}
(\textit{init}). Assuming the duration of $\mathcal{HT}$ is finite, 
there must exist some distinguishable last completion event (\textit{comp}). 
%Operationally, an execution of $\mathcal{HT}$ begins at \textit{init}. 
After a transaction step $\varphi$ is executed, the set of concurrent state
update $\Gamma_{\varphi}$ is carried out and the value of boolean predicates
in $\Delta_{\varphi}$ is evaluated to determine the next step
$\varphi_{next}$, that is, a predicate $\beta_i \in \Delta_{\varphi}$
enables a transaction step.  This process continues until $\mathcal{HT}$
reaches \textit{comp} state.  Note that $\mathcal{FT}$ can update some state
elements of $\mathcal{HT}$ during its write action, based on which
predicates in $\Delta_{\varphi}$ may be computed. 
Figure~\ref{fig:interleaving2} gives a fragment of $\mathcal{HT}$
corresponding to the top module of UART IP.  Let us consider a possible
execution of $\mathcal{HT}$, $(init \rightarrow t_1 \rightarrow t_4
\rightarrow t_6 \rightarrow comp)$.  At $init$, the FW sets the HW input
signal, $(rst\_i==0)$, based on which the execution follows the edge to
$t_1$.  At $t_1$, the predicate, $cs$, is computed based on HW input
signals, $cyc\_i$, $stb\_i$, $adr\_i$ and $we\_i$.  Assuming that $cs\_i
\&\& we\_i$ holds true, the execution proceeds to step $t_4$ which computes
predicates based on the values in $adr\_i$.  The next step is determined by
the values of predicates, $'UART\_MC, 'UART\_IE, 'UART\_CT$.  Assuming that
$'UART\_MC$ holds true, the execution proceeds to $t_6$.  During the
execution of $t_6$, three parallel state update $(dtr\_no, rts\_no,
loopback)$ and two predicate evaluation ($\beta_1,\beta_2)$ is performed
based on the updates to state elements.  After $t_6$, the execution of
$\mathcal{HT}$ leads to $comp$ and the transaction finishes its execution. 
Note that the dotted edges in $\mathcal{HT}$ denotes that eventually the
execution reaches $comp$.  The update of state elements are shown in
rectangular boxes, whereas the evaluation of boolean predicates are shown
inside the rounded boxes.  The blue arrows point to the state updates or
predicate evaluation at a particular~$\varphi$.

\begin{figure}[t]
\scalebox{.55}{\import{figures/}{fw-hw-interleaving.pspdftex}}
\caption{Some HW Transactions for UART
\label{fig:interleaving2}}
\end{figure}

\textbf{Scenario/Transaction Pairs}
%
A scenario, $\mathcal{S}$, restricts the interaction between $\mathcal{HT}$
and $\mathcal{FT}$.  Similar to service-function based transaction in
TLM~\cite{hvc}, we view $\mathcal{FT}$ and $\mathcal{HT}$ as a set of
service in response to its interacting $\mathcal{HT}$ or $\mathcal{FT}$,
respectively.  A {\em Scenario/Transaction} pair is a set of valid
interactions between $\mathcal{HT}$ and $\mathcal{FT}$ for a
particular~$\mathcal{S}$.  For example, an UART IP contains a {\em
transmitter} and a {\em receiver} module.  Let us assume that a scenario,
$\mathcal{S}$, configure a UART in {\em transmission} mode.  Then the
interaction between $\mathcal{FT}$ and $\mathcal{HT}$ is constrained to
contain only those interactions that are relevant to the {\em transmitter}
module.  By contrast, the interactions with respect to the {\em receiver}
module is irrelevant for $\mathcal{S}$.  However, when $\mathcal{S}$
configures a UART in {\em loopback} mode, then the interaction between
$\mathcal{FT}$ and $\mathcal{HT}$ must contain transactions in {\em
transmitter} as well as {\em receiver} module.

This constrained interactions can be exploited 
to decompose the verification state-space. 
This requires an automatic technique for scenario driven 
partitioning of the FW-HW transactions to generate only 
relevant {\em Scenario/Transaction} pairs. Conventional 
approaches for co-verification construct a unified 
FW-HW model in FSM or BPDS that restrict such decomposition. 
In this paper, we built a unified FW-HW model in C program 
which enable us to automatically generate such {\em scenario/transaction} 
pairs using path-based symbolic execution techniques. The details 
of our technique are explained in the next section.
%
\Omit{
\textbf{Problem Statement}
Given a FW model in C, HW models in Verilog RTL, a property $\psi$ and 
a set of scenarios, the goal is to check whether $\psi$ holds under all 
scenarios in the combined FW-HW model. 
}
%
\section{Proposed Methodology}
Figure~\ref{proposed-flow} describes our overall tool framework. In the
subsequent section, we describe each step in details.

>>>>

%
%===============================================================================
\section{Introduction}
%===============================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Pre-RTL versus Post-RTL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-RTL and post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (IP), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers/firmware) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or firmware).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing.

In the post-RTL phase, when the RTL design is available, hardware/software
co-verification becomes more complex.  Unlike the TLM, which only captures
limited design functionality, the RTL implements cycle-accurate behavior
of the final HW, and addresses concerns related to power, area and
timing.  Verification is mandatory to assert that the driver or firmware (FW) 
is executed correctly together with the RTL.  The effectiveness of testing 
is thus severely limited in post-RTL co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Challenges in Co-verification 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
However, the verification of FW given in C/C++ together with the RTL given in a hardware
description language (HDL) such as Verilog/VHDL is a challenging task due to 
few reasons -- 1) the timing and semantic mismatch between the synchronous
clock-driven HW model and asynchronous event-driven SW model -- for example,
FW running on a processor could be much faster than the HW
model it interacts with, 2) lack of co-specification model that precisely
captures the design intent of the interaction behavior, 3) lack of automated
formal tools that support co-designs that are mostly implemented in C and 
Verilog/VHDL.   In this paper, we address some of these issues to facilitate 
wider adoption of formal techniques in hardware/software co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Handling Semantic Gap  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Generating a unified FW-HW model~\cite{fmsd02,emsoft07,fase10} for 
co-verification is a well-known phenomenon to handle the semantic mismatch 
between the HW~\footnote{Note that whenever we refer to ``hardware'' from this point onwards, we mean an RTL implementation 
and not a TLM.}  and FW models.  Most of the existing techniques are not 
useful in practise for three reasons -- 1) use of abstract HW models 
instead of register-transfer level implementation, 2) the HW-FW model
construction is agnostic to the underlying reasoning engine, 3) does not
leverage the co-specification model, hence ineffective for practical
co-verification.  We will call the unified FW-HW model as \emph{co-verification model}.

In this paper, we present two different approaches for automatic formal
verification of HW/SW co-designs -- \emph{monolithic} and \emph{path-based}.  
We embody these techniques in our tools, \hwcbmcv and \verifox, respectively. 
A desirable characteristic of a co-verification model is the ability to exploit 
the expressivity of the underlying reasoning engines, for example, constructing 
a bit-level co-verification model for SAT solvers or a word-level model for SMT 
solvers.  \hwcbmcv constructs a co-verification model through 
in tandem symbolic execution of the HW and SW models.  To this end, a bit-level 
netlist (typically represented as And Inverter Graph) or word-level netlist 
(represented in tool specific format) is synthesized from a HW Register Transfer 
Level (RTL) implementation. Similarly, the SW, given in C, is automatically 
translated into a Static Single Assignment (SSA) form.  A monolithic verification 
condition is generated from these two models which is then solved using a 
satisfiability solver.  Whereas, a co-verification model in \verifox is constructed 
by first translating the RTL description into a software representation in C followed 
by sequentializing the hardware and firmware models.  This results in a single
threaded sequential co-verification model which is represented in C. A path-wise 
symbolic execution is then performed over the co-verification model to
effectively mine only relevant interaction behaviors between the SW and HW
models.

%A key to building an effective co-verification tool is to translate the HW
%and SW models to a common formal semantics called, 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Co-specification Model  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A co-verification models must leverage the specification of system-level 
properties across HW/SW boundaries for uncovering critical co-design bugs 
and for effective co-verification.  Our co-specification models are 
specified in C with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. Our
co-verification models in \hwcbmcv and \verifox can be easily verified against 
these co-specification models. The co-design properties expressed by 
co-specification models are sufficiently expressive, amenable to 
verification reuse and can be readily used in assertion-based verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %       Scenario, Transaction and Scenario/Transaction Pair
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A firmware only exercise some fragment of hardware state-space in a typical 
hardware/software co-design~\cite{dac17,taa,ibm-other}.  
A firmware services the requests from the application software 
layer or its interacting hardware.  Similarly, the hardware services 
the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in 
\emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the 
UART in \emph{transmitter} mode which restricts the communication between the FW 
and HW to set of services that only transfers data from the FW to the serial output 
port of the transmitter HW while completely bypassing the receiver HW.  
A {\em Scenario/Transaction} pair is a set of valid interactions between 
$\mathcal{HT}$ and $\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable 
and effective co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %        Difference between Monolithic and Path-wise Approach
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A monolithic approach in \hwcbmcv based on Bounded Model Checking~\cite{CKY03}, 
cannot exploit this infeasibility.  Whereas, a path-based symbolic execution in 
\verifox can automatically infer such Scenario/Transaction pairs and check their 
validity using backend solvers.  For building effective and scalable formal 
co-verification tool, it is therefore imperative to automatically mine only 
relevant scenario/transaction pairs that occurs in practical co-designs.

\Omit{
A system is comprised of a set of concurrent FW and HW transactions which
interleaves asynchronously. It is well recognized that concurrency is at
the heart of co-verification. The number of interleavings between the FW
and HW threads can be large, which can become a bottleneck in the
verification process.

The concurrency between the HW and SW components is a key challenge in
co-verification.  In this paper, we exploit specific interaction patterns
between the FW and HW to reduce the number of interleavings.  We observe
that this interaction pattern follows a producer-consumer relationship --
the master module (producer) solely updates the shared registers which are
consumed by the slave (consumer).  This key observation allows us to
sequentialize a concurrent FW-HW model following specific interaction
patterns.
}

@@

\end{enumerate}
%
%===============================================================================
\section{Working Example}
%{\bf Running example: }
%===============================================================================
In this section, we present an working example of FW and HW model to explain our 
co-verification methodologies.

\textbf{Firmware Model} Figure~\ref{firmware} gives a fragment of a FW model for a Universal 
Asynchronous Transmitter/ Receiver (UART) IP.  The
$wb$ class of functions communicate with the {\em wishbone} bus interface to
set/reset the UART input ports and read/write data through the bus
interface.  The FW implements Linux style inb() and outb() functions which
invokes the $wb$ functions to communicate with the UART.  The FW begins 
by resetting the UART IP which is followed by a $wb\_idle()$ function.  The 
FW then configures the UART in {\em loopback} mode using $outb$ function calls. 
Inside the loop, the firmware checks for a {\em transmitter} or {\em
receiver} interrupt by reading the Interrupt Status register $(UART\_IS)$. 
Depending on the type of interrupt, the firmware either receives a data packet 
or sends a data packet over the serial output.

\rmcmt{The $set\_inputs()$ and
$next\_timeframe()$ functions, used in $wb$ functions, are 
primitives in \textsc{HW-CBMC} that direct the tool to set the 
inputs to the HW signals and advance the HW clock, respectively.}

\textbf{Hardware Model} Figure~\ref{fig:interleaving2} shows a fragment of the state machine 
implementation of the UART hardware RTL.  The UART IP contains a
\emph{transmitter} module, a \emph{receiver} module and a \emph{clock
generator}.  The update of state elements (\emph{registers}) and 
\emph{wires} are shown in rectangular boxes and rounded boxes respectively.

\textbf{Co-specification Model} We verify that the transmitted data is the same as the 
received data in the {\em loopback} mode.  To do so, we annotate the FW model 
with an assertion $assert()$, as shown in Figure~\ref{firmware}.  Note that the 
array containers $rx\_b$ and $tx\_b$ stores the data obtained through the $inb$
and $outb$ function calls which directly reads and writes the UART IP ports
respectively. 
\[
  for(i=0; i<=N; i++) 
   \quad assert(rx\_b[i] == tx\_b[i]);
\]
%
We formally describe the notion of \emph{transaction} and \emph{scenario} with the
help of our working example. \\
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l}
\hline
Wishbone Interface & Main Module \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
typedef unsigned char u8;
unsigned char inb 
 (unsigned long port) {
  return wb_read(port);
}
void outb (u8 value, 
 unsigned long port) {
  wb_write(port, value);
}
void wb_reset(void) {
 rst_i = 1;
 set_inputs();
 next_timeframe();
 rst_i = 0;
 stb_i = 0; 
 cyc_i = 0;
}
void wb_idle() {
 set_inputs();
 next_timeframe();
}
void wb_write(_u32 addr, 
     _u8 b) {
 adr_i = addr;
 dat_i = b;
 we_i = 1;
 cyc_i = 1;
 stb_i = 1;
 set_inputs();
 next_timeframe();
 we_i = 0;
 cyc_i = 0;
 stb_i = 0;
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
int main() {
wb_reset();
wb_idle();
// Configure the uart
outb (0x13, UART_MC);  
outb (0x80, UART_CM3);
outb (0x00, UART_CM2);
outb (0x00, UART_CM1);
outb (0x00, UART_CR);  
outb (0x03, UART_IE); 
char tx_b[] = "Hello world";
_u8 status = 0;
char rx_b[100];
int i=0,c=0,d=0;
// data transfer in loopback
for (i=0; i<1990; i++){
if (irq_o){
 status=inb(UART_IS)&0x0c;
 if(istatus==0x0c){
 //it was a tx_empty interrupt
 outb(*(tx_b+c),UART_TR); c++;
 }
 else{ //status==0x04
 //it was an rx_data interrupt
  rx_b[d] = inb(UART_TR); d++;
 }
}else {
 // no interrupt. 
 wb_idle();
 wb_idle();
 }
}
// property
for(i=0; i<=10; i++)
 assert(rx_b[i] == tx_b[i]);
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Firmware model of UART IP}
\label{firmware} 
\end{figure}
%
\textbf{Scenario ($\mathcal{S}$):}
Let $B$ denote a set of all possible behaviors of a system. A scenario is a set of 
constraints $\mathcal{S}$ that restrict the system behaviors
to $B' \subseteq B$. For example, the left hand side of Figure~\ref{fig:st} 
gives an example scenario for the UART FW shown in Figure~\ref{firmware}. 
Here, $\mathcal{S}$ configure the UART IP in {\em loopback} mode with 
interrupt enabled.

@@

Each execution of $\mathcal{HT}$ is a non-empty 
sequence of transaction steps where each transaction step 
happens in a single clock. The first step is called {\em initiation event}
(\textit{init}). Assuming the duration of $\mathcal{HT}$ is finite, 
there must exist some distinguishable last completion event (\textit{comp}). 
%Operationally, an execution of $\mathcal{HT}$ begins at \textit{init}. 
After a transaction step $\varphi$ is executed, the set of concurrent state
update $\Gamma_{\varphi}$ is carried out and the value of boolean predicates
in $\Delta_{\varphi}$ is evaluated to determine the next step
$\varphi_{next}$, that is, a predicate $\beta_i \in \Delta_{\varphi}$
enables a transaction step.  This process continues until $\mathcal{HT}$
reaches \textit{comp} state.  Note that $\mathcal{FT}$ can update some state
elements of $\mathcal{HT}$ during its write action, based on which
predicates in $\Delta_{\varphi}$ may be computed.

Figure~\ref{fig:interleaving2} shows a fragment of $\mathcal{HT}$
corresponding to the top module of UART IP.  Let us consider a possible
execution of $\mathcal{HT}$, $(init \rightarrow t_1 \rightarrow t_4
\rightarrow t_6 \rightarrow comp)$.  At $init$, the FW sets the HW input
signal, $(rst\_i==0)$, based on which the execution follows the edge to
$t_1$.  At $t_1$, the predicate, $cs$, is computed based on HW input
signals, $cyc\_i$, $stb\_i$, $adr\_i$ and $we\_i$.  Assuming that $cs\_i
\&\& we\_i$ holds true, the execution proceeds to step $t_4$ which computes
predicates based on the values in $adr\_i$.  The next step is determined by
the values of predicates, $'UART\_MC, 'UART\_IE, 'UART\_CT$.  Assuming that
$'UART\_MC$ holds true, the execution proceeds to $t_6$.  During the
execution of $t_6$, three parallel state update $(dtr\_no, rts\_no,
loopback)$ and two predicate evaluation ($\beta_1,\beta_2)$ is performed
based on the updates to state elements.  After $t_6$, the execution of
$\mathcal{HT}$ leads to $comp$ and the transaction finishes its execution. 
Note that the dotted edges in $\mathcal{HT}$ denotes that eventually the
execution reaches $comp$.  The update of state elements are shown in
rectangular boxes, whereas the evaluation of boolean predicates are shown
inside the rounded boxes.  The blue arrows point to the state updates or
predicate evaluation at a particular~$\varphi$. \\
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{fw-hw-interleaving.pspdftex}}
\caption{Some HW Transactions for UART
\label{fig:interleaving2}}
\end{figure}
%
\textbf{Scenario/Transaction Pairs}
%
A scenario, $\mathcal{S}$, restricts the interaction between $\mathcal{HT}$
and $\mathcal{FT}$.  Similar to service-function based transaction in
TLM~\cite{hvc}, we view $\mathcal{FT}$ and $\mathcal{HT}$ as a set of
service in response to its interacting $\mathcal{HT}$ or $\mathcal{FT}$,
respectively.  A {\em Scenario/Transaction} pair is a set of valid
interactions between $\mathcal{HT}$ and $\mathcal{FT}$ for a
particular~$\mathcal{S}$.  For example, an UART IP contains a {\em
transmitter} and a {\em receiver} module.  Let us assume that a scenario,
$\mathcal{S}$, configure a UART in {\em transmission} mode.  Then the
interaction between $\mathcal{FT}$ and $\mathcal{HT}$ is constrained to
contain only those interactions that are relevant to the {\em transmitter}
module.  By contrast, the interactions with respect to the {\em receiver}
module is irrelevant for $\mathcal{S}$.  However, when $\mathcal{S}$
configures a UART in {\em loopback} mode, then the interaction between
$\mathcal{FT}$ and $\mathcal{HT}$ must contain transactions in {\em
transmitter} as well as {\em receiver} module.

\rmcmt{This constrained interactions can be exploited 
to decompose the verification state-space. 
This requires an automatic technique for scenario driven 
partitioning of the FW-HW transactions to generate only 
relevant {\em Scenario/Transaction} pairs. Conventional 
approaches for co-verification construct a unified 
FW-HW model in FSM or BPDS that restrict such decomposition. 
In this paper, we built a unified FW-HW model in C program 
which enable us to automatically generate such {\em scenario/transaction} 
pairs using path-based symbolic execution techniques. The details 
of our technique are explained in the next section.}
%
\Omit{
\textbf{Problem Statement}
Given a FW model in C, HW models in Verilog RTL, a property $\psi$ and 
a set of scenarios, the goal is to check whether $\psi$ holds under all 
scenarios in the combined FW-HW model. 
}
%
\section{Proposed Methodology}

\subsection{Monolithic Co-verification Approach}

\subsection{Path-based Co-verification Approach}
%
Figure~\ref{proposed-flow} describes our overall tool framework. In the
subsequent section, we describe each step in details.
